---
title: "Create Beamer Slide with R Markdown"
subtitle: "A Reproducible Example"
author: "Yongfu, Liao"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
institute: "National Taiwan University"
fontsize: 9pt
output:
  beamer_presentation:
    theme: "Madrid"
    colortheme: "beaver"
    fonttheme: "structurebold"
    incremental: true
    latex_engine: xelatex
    includes:
        in_header: header_beamer.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
date_text <- "format(Sys.Date(), '%B %d, %Y')"
```


## Reproduce This slide: yaml Header

```yaml
---
title: "Create Beamer Slide with R Markdown"
subtitle: "A Reproducible Example"
author: "Yongfu, Liao"
date: "`r date_text`"
institute: "National Taiwan University"
```

## Reproduce This slide: yaml Header (Cont.)
```yaml
fontsize: 9pt
output:
  beamer_presentation:
    theme: "Madrid"
    colortheme: "beaver"
    fonttheme: "structurebold"
    incremental: true
    latex_engine: xelatex
    includes:
        in_header: header_beamer.tex
---
```

> - Links  
[header_beamer.tex](https://github.com/liao961120/local_depend/blob/master/style_sheets/header_beamer.tex)  
[Source Dir](https://github.com/liao961120/local_depend/tree/master/GitHubPage/my_R_book/Rmarkdown_features/beamer)


## Two-Column Layout

\begin{columns}
\column{0.75\textwidth}

```{r out.width="75%"}
library(ggplot2)
ggplot(iris, aes(x = Sepal.Width, y = Sepal.Length)) +
    geom_point(aes(color=Species), show.legend = F)
```

\column{0.25\textwidth}
Some Text Some Text Some Text Some Text Some Text Some Text Some Text Some Text Some Text.
\end{columns}

## Two-Column Layout: Code

```tex
\begin{columns}
\column{0.75\textwidth}
```
    `r ''````{r out.width="75%"}
    library(ggplot2)
    ggplot(iris, aes(x = Sepal.Width,
                     y = Sepal.Length)) +
        geom_point(aes(color = Species),
                   show.legend = F)
    ```
```tex
\column{0.25\textwidth}
Some Text Some Text Some Text Some Text Some Text Some Text Some Text Some Text Some Text.
\end{columns}
```

## Various Blocks
 
\begin{alertblock}{Red Block}
$\frac{\int_{-\infty}^{\infty} f(x)dx}{b}$
\end{alertblock}
 
\begin{examples}
Example Blocks are green.
\end{examples}

### Blue Block
Paragraph 1

Paragraph 2

## Various Blocks: Code

**Red Block**: Can set titles
```tex
\begin{alertblock}{Red Block}
$\frac{\int_{-\infty}^{\infty} f(x)dx}{b}$
\end{alertblock}
```

**Examples**: Can't change title
```tex
\begin{examples}
Example Blocks are green.
\end{examples}
```

**Blue Block**: h3 level Headings[^h3]
```markdown
### Blue Block
Paragraph 1: Text In Box

Paragraph 2: Blahblahblah
```

[^h3]: Note this only works in 'Madrid' theme in R markdown.


## Incremental Bullets

Remember to set `incremental: true` in the yaml header if you want this feature.

1. Item 1

2. Item 2

3. Item 3
