---
title: "Multivariate Analysis R Notes"
author: Yongfu, Liao
date: March 24, 2018
output:
    html_document:
        css: /Users/user/local_depend/style_sheets/style.css
        code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE,
	dev = 'svg',
	fig.height = 3, 
	comment = ""
)
library(dplyr)
library(latex2exp)
library(ggplot2)
theme <- theme(axis.text.x = element_text(size = 7, face = "plain", angle = 30),
               axis.text.y = element_text(size = 7, face = "plain"),
               axis.title.x = element_text(size = 9, face = "bold"),
    axis.title.y = element_text(size = 9, face = "bold"))
```

Bivariate Normal Contour Map
-----------------------------------

### `ellipse` function
```r
ellipse(x, scale, centre, level, npoints = 1000)
```
- `x`: a single number, correlation of the two variables.

- `scale`: vector, **standard deviation** of the two variables.

- `centre`: vector, center of the ellipse, i.e. the mean vector of the bivariate normal distribution.

- `level`: a single number, the contour probability.

- `npoints`: number of points used to draw the contour.

`ellipse` returns a **matrix** with dimension (`npoints` $\times$ 2), which can be used to plot contour.

### Data Generation
The `for` loop below is used to generate a data frame with 3 columns(variables):
    - Column 1: First variable of bivariate normal function ($x_1$)
    - Column 2: Second variable of bivariate normal function ($x_2$)
    - Column 3: The contour that $x_1$ & $x_2$ on the same row belongs to.
```{r}
library(ellipse)

All_contours <- c(NA, NA, NA) 
    ## Set empty start for appending ##

for (i in 1:5) {
    level <- 0.1*i 
        ## Set Contour prob., prob. of obs within contour ##
    ell_data <-ellipse(-0.8, c(sqrt(2), 1), centre = c(1, 3), level = level, npoints = 800+(i-1)^3)
        ## npoints: bigger contours with more points ##
    class <- rep(paste(level*100, "% Contour", sep=""), nrow(ell_data))
        ## Assign contour class ##
    ell_data <- as.data.frame(ell_data)
        ## Change to data.frame BEFORE cbind, ##
        ## or coersion happens ##
    ell_data <- cbind(ell_data, class)
    
    All_contours <- rbind(All_contours, ell_data)
}

All_contours <- All_contours[-1,]
    ## Remove the empty start ##
```

### Plotting
```{r}
ggplot(data = All_contours) +
    geom_point(aes(x = x, y = y, color = class),
               size = 0.1) +
    scale_colour_grey(start = 0.7, end = 0.3) +
        ## Use gray scales instead of colored default ##
    labs(color = "Contours", 
         title = "Contour Plot",
         x = TeX("$x_1$"), y = TeX("$x_2$")
    )
```

Multivariate Normal Functions
----------------------------------

### Generate density f(x)

```{r echo=TRUE}
library(mvtnorm)

mu <- c(1, 3) # mean vector
Sigma <- matrix(c(2, -0.8*sqrt(2), -0.8*sqrt(2), 1),
                nrow = 2) # covariance matrix

dmvnorm(x = c(2, 5), mean = mu, sigma = Sigma)
```

- `x`: Vector x in f(x), all variables of the multivariate normal distribution.
- `mean`: Mean vector(center of ellipse) of the multivariate normal distribution.
- `sigma`: Covariance matrix of the multivariate normal distribution.

`dmvnorm` returns f(x), the range of the multivariate normal function. For example, `dmvnorm(x = c(2, 5), mean = mu, sigma = Sigma)` returns the value f($x_1=2$, $x_2=5$) of the multivariate normal distribution specified by mean vector, `mu`, and covariance matrix, `Sigma`.

#### Example: Densities of a Contour 
```{r echo=TRUE}
data <- All_contours %>% 
    filter(class == "50% Contour")

dmvnorm(x = data[1, 1:2], mean = mu, sigma = Sigma)[[1]]
dmvnorm(x = data[4, 1:2], mean = mu, sigma = Sigma)[[1]]
```

The retured values are the same(very close), since they are on the same contour. See the section [above]{#data-generation} for more details.

### Covariance Matrix

Generater covariance and correlation Matricies:
```{r echo=TRUE}
library(mat2tex)
cov.mt <- cov(iris[,1:4]) ## Cov Matrix of variable 1~4
cor.mt <- cor(iris[,1:4]) ## Cor Matrix of variable 1~4
```

Covariance matrix 
```{r echo=FALSE, results='asis'}
"$=" %_% xm(cov.mt,2) %_%"$"
```

Correlation matrix 
```{r echo=FALSE, results='asis'}
"$=" %_% xm(cor.mt,2) %_%"$"
```


