---
title: 'Q5: Food Price'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message=FALSE,
                      warning=FALSE,
                      dev='svg',
                      fig.dim=c(3.6, 3.2))
library(readr)
library(tidyverse)
library(knitr)
library(kableExtra)
```

```{r data_import}
fp <- read_table2("data/FOODP.txt", col_names = FALSE, 
    skip = 5)[1:24, -1]
colnames(fp) <- c("city", paste("food",1:5,sep=""))
```

(a) Scree Plot, PCs, Interpretation
=======================================
```{r scree_plot}
library(factoextra) # For Visualizing MvA
PCA <- prcomp(fp[,-1], scale=T)
# fviz_eig(PCA)

eigva <- as_data_frame(PCA[["sdev"]]^2)
scree <- cbind(1:5 , eigva)
scree <- cbind(scree, "a")
colnames(scree) <- c("PC", "var", "a")

ggplot(scree, aes(x = PC, y = var, group = a)) +
    geom_point() + geom_line() +
    scale_y_continuous(
        sec.axis = sec_axis(~ . / sum(eigva) * 100,
                            name = "Variance Accounted (%)")
                   )+
    labs(x="Principle Components",
         y=latex2exp::TeX("$\\hat{\\lambda_i}$"),
         title="Scree Plot")    
```

By either the "elbow" of the scree plot or the unity criterion, **one principle component** is sufficient, which explains **`r round(eigva[1,1]/sum(eigva)*100, 2)`%** of the total variance.

```{r PC_loadings}
pc <- PCA[["rotation"]]
kable(round(pc,3),"markdown", align = "c")
```

(b) The First Two principal component scores
===================================================

```{r}
# PCA[["x"]] returns the "rotated" score(pc score) of the original data
pc_score12 <- as.data.frame(PCA[["x"]][,1:2])
pc_score12 <- cbind(fp$city, pc_score12) %>% rename(city=`fp$city`)
kable(pc_score12, "markdown", align="c",digits = 2)
```


(c) Scatter Plot: PC1 vs. PC2
============================================
```{r fig.dim=c(7, 4)}
ggplot(pc_score12, aes(x=PC1, y=PC2))+
    geom_point()+
    geom_text(aes(label=city),
              hjust=-0.1, vjust=0, size = 3)+
    scale_x_continuous(limits = c(-4.6, 2.9), breaks = seq(-4.5,3, 0.5))+
    scale_y_continuous(limits = c(-1.7, 1.5), breaks = seq(-1.75,1.5, 0.25))+
    labs(title="Scatter Plot of PC1 & PC2",
         x="Principle Component 1",
         y="Principle Component 2")
```

(d) Food Price in Different Cities
===============================================
```{r}
## Max PC score: cheapest food price
cheap <- pc_score12$city[pc_score12$PC1 == max(pc_score12$PC1)]

## min PC score
expense <- pc_score12$city[pc_score12$PC1 == min(pc_score12$PC1)]
```

Using principle component 1 as indicator, since it has the same sign on all "weighted" food price, and it's the most important principle component in explaining the variability of the data, the x-axis of the scatter above showed that **`r cheap`** has the cheapest food price, and **`r expense`** has the most expensive food price.