---
title: "Advanced Graphics"
output:
  html_document:
    toc: true
    toc_float: true
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.path = "Advanced Graphics/")
```
```{r, include=FALSE}
library(ggplot2)
library(dplyr)
```

## Six Principles of Graphing
1. Comparison

2. Mechanism

3. Multivariate Data

4. Integrating evidence: <br/>Different forms of expressions

5. Credibility: <br/>Describing and documenting the evidence with sources       appropriate labels, and scales.

6. Content is King: <br/>Garbage in, garbage out.

## Exploratory Graphs 
1. DON'T use exploratory graphs to communicate results.

2. Initial step in an investigation:<br/>Used to point the data scientist in a fruitful direction.

## ggplot2
### Grammar(template) of ggplot2
```{r, eval=FALSE}
ggplot(data = <DATA>) + 
  <GEOM_FUNCTION>(
     mapping = aes(<MAPPINGS>),
     stat = <STAT>, 
     position = <POSITION>
  ) +
  <COORDINATE_FUNCTION> +
  <FACET_FUNCTION>
```

### Aestheics
An aesthetic is a **visual property** of the objects in the plot. Aesthetics include things like the size, the shape, the color, and transparency of the points.

Each `geom_` function has its suitable aesthetics, which can be found by typing `?geom_name` in the console.

`alpha` sets transparency.

`color` adds color to points and lines, color boarder to bars.

`fill` fills color in histogram or bar chart.

### Coordinate Functions
```r
coord_cartesian(xlim=c(x_min,x_max), ylim=c(y_min,y_max))
```
sets the range of x and y-axis.

### Example: box plot
```{r, echo=TRUE}
ggplot(data = mpg) +
    geom_boxplot(
      mapping = aes(x = reorder(class, hwy, FUN = median), 
                    y = hwy)) +
    coord_flip()  # 將x, y軸對調
```

### Example: geom_smooth
```{r}
ggplot(diamonds, 
aes(x=carat, 
    y=price, 
    color=clarity)) +
geom_smooth(method="lm", formula=y~poly(x, 2))

ggplot(diamonds, 
aes(x=carat, 
    y=price, 
    color=clarity)) +
geom_smooth(method="lm", formula=y~poly(x, 5))
```

### Example: lable on grapgh
```{r}
ggplot(mtcars, aes(x=wt, y=mpg)) +  geom_label(label=rownames(mtcars), nudge_y=0.5) + geom_point()
```
`nudge_y = `, `nudge_x = ` adjust the position of the lable.


## Interactive Graphic

```{r}
library(plotly)
gp=ggplot(mtcars, aes(x=wt, y=mpg)) +
  geom_point(aes(text=rownames(mtcars)))
ggplotly(gp)
```
**plotly** is compatible with **ggplot2**.
