---
title: "Rmarkdown Features"
subtitle: "This is subtitle"
author: Yongfu Liao
date: March 18, 2018
bibliography: ./citation/neanderthals.bib
csl: ./citation/apa-no-doi-no-issue.csl
link-citations: yes
output:
  html_document:
    css: ../style_sheets/style.css
    df_print: kable
    code_folding: show
---
<!--more-->

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

## yaml Setup {.tabset}

### Citation

Include the code in yaml header.

```{yaml}
bibliography: ./citation/neanderthals.bib
csl: ./citation/apa-no-doi-no-issue.csl
link-citations: yes
```

### HTML Output Options

Include the code in yaml header.
```{yaml}
output:
  html_document:
    df_print: kable
    code_folding: hide
    fig_width: 7
    fig_height: 6
```


## Dynamic Variables

Include **R Code Chunks' Variables** in the body of text.

```{r}
alpha <- 1:3
text <- c("a", "b")
```


1. The numeric variable $\alpha =$ `r alpha[1]`.
    ```{yaml}
    # The numeric variable $\alpha =$ `r alpha[1]`.
    ```

2. The string variable $text =$ `r text[1]`.
    ```{yaml}
    # The string variable $text =$ `r text[1]`.
    ```

## $\LaTeX$ Rendering {.tabset}

### kable
```{r}
library(knitr)
library(dplyr)
iris_data <- head(iris) %>% select(Sepal.Length, Sepal.Width)

colnames(iris_data) <- c("$\\alpha$", "$\\frac{a}{b}$")
kable(iris_data, format = "markdown", align = "c")
```

### Plotting Expression
```{r}
library(latex2exp)
library(ggplot2)
iris_data2 <- iris[c(1:3, 50:53, 100 :103),] %>% select(Sepal.Length, Sepal.Width, Species)

iris_data2 %>% ggplot() +
    geom_point(mapping = aes(x=Sepal.Length, y=Sepal.Width, color=Species))+
    labs(x=TeX("Length($\\frac{kg}{cm^2}$)"),
         y=TeX("\\frac{a}{b}"))
```



## Citations 
To include **Citations**[^1], [Read this](https://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html){target="_blank"}.

The text cited [@RN9].

@RN15 主張...

### Include `.bib` file in Reference {.tabset}

#### Specific References
Include specific references, which are **not cited** in the body of text.
```{pandoc}
---
nocite: |
  @item1, @item2
...
```

#### All References
Include **All references** in the body of text.
```{pandoc}
---
nocite: |
  @*
...
```

## Emphasis {.tabset}

### Full Example
```{html}
<div class="success">
| This is pragraph 1. [**Red text**]{.red}, line 1. The text[^2] is abcd 中文
| This is [line two]{style="color:rgba(255, 0, 0, 1)"}. 

Paragraph 2 starts here.
</div>
```

<div class="success">
| This is [**Red text**]{.red}, line 1. The text[^2] is abcd 中文
| This is [line two]{style="color:rgba(255, 0, 0, 1)"}. 

Paragraph.
</div>

### Colored divs
There are 4 choices of class: 
1. success
2. info
3. warning
4. error
```{html}
<div class="success">
    some text in the colored region
</div>
```





[^1]: This is footnote 1.

[^2]: This is footnote 2.

    long footnotes.
 
<!-- syntax of pandoc: reference will be added here --> 
# References 